\storeglosentry{gls:activerecord}{
  name={active record}, description={Design pattern describing a solution to the
  problem of programmatic interaction with data in a database. In active record
  interaction is handled seamlessly by methods, such as save and delete, on
  objects that corresponds to entries in the database.}}


\storeglosentry{gls:anti_patterns}{
  name={anti-patterns}, 
  description={An ineffective practice commonly used.}
}

\storeglosentry{gls:appstate}{
  name={application state}, description={REST sees a Web application as a state
  machine that transitions to different states by following links and submitting
  forms in hypermedia. Application state is the current state of the application
  given by the client's sequence of followed links and submitted forms.}  }

\storeglosentry{gls:business_logic}{
  name={business logic}, description={The part of the application that defines
  the data model, the data storage, the data retrieval, and functions that
  retrieve certain data from the models.}}

\storeglosentry{gls:cgi}{
  name={CGI}, 
  description={A standard that connects a Web server with any programming language.
  } 
}

\storeglosentry{gls:cloud}{
  name={cloud}, 
  description={A term for the Internet often used when
  describing services on the Internet where most of the infrastructure is
  hidden. An example is Google App Engine that hides all server details from its
  users. } 
}

\storeglosentry{gls:qa}{
  name={quality attributes}, 
  description=A measurable physical or abstract property of quality.
}


\storeglosentry{gls:requestresponsepattern}{
  name={request-response pattern},
  description={Architectural pattern for message exchange where every request is matched with a response.}
}


\storeglosentry{gls:base_32}{
  name={Base32}, 
  description={An alphabet based on 32 symbols.}
}

\storeglosentry{gls:dom}{
  name={DOM}, description={Tree representation format of XML documents, often used from
  JavaScript to access and modify a HTML document.}  
}

\storeglosentry{gls:content_negotiation}{
  name={content negotiation}, description=HTTP mechanism that serves a certain
  representation of a resource based on the capabilities of the client
  \citep[Sec. 12]{w3:HTTP} 
}

\storeglosentry{gls:fifo}{
  name={First In, First Out}, 
  description=A way of storing or processing of data where the first input goes out first.
}

\storeglosentry{gls:geocoding}{
  name={geocoding}, 
  description=Process of converting a location name into a latitude and longitude pair.
}

\storeglosentry{gls:googleaccounts}{
  name={Google accounts},
  description={Service that provides users access to several Google services.}
}

\storeglosentry{gls:location_based_service}{
  name={location-based service},
  description=A service that takes the location of the user into consideration.
}

\storeglosentry{gls:meridian}{
  name={meridian}, 
  description={A half great circle linking the North Pole and the South Pole.}
}

\storeglosentry{gls:pagination}{
  name={pagination}, 
  description=Process of splitting up entities to several Web pages.
}

\storeglosentry{gls:ttl}{name={time to live}, description=The period of time some data is valid.}

\storeglosentry{gls:wsgi}{name={WSGI}, description=A specification for the
interface between Web servers and Python Web application frameworks. The
rationale of the specification is to decouple the Web application framework from
the Web server.}

\storeglosentry{gls:wsdl}{name={WSDL}, description={An XML language to describe
the interface of a Big Web Service.}}%
%
\storeglosentry{gls:soap}{
  name={SOAP}, 
  description={A wrapper around XML documents used when transferring XML documents.}}%
%
\storeglosentry{gls:ws}{name={WS-* stack}, description={A set of
Big Web Service descriptions.}}%
%
\storeglosentry{gls:bigws}{name={Big Web Services}, description={Term coined in
                           \citep[ch.10]{rest:webservice}, for Web Services
                           architecture based on SOAP, WSDL, and the WS-*
                           stack.}}%
%
\storeglosentry{gls:disthyp}{name={distributed hypermedia}, description={A
           non-linear media that integrates other media besides text, that can be
           located at different places.}}%
%
\storeglosentry{gls:uri}{
  name={URI}, description={Acronym for Uniform Resource Identifier: scheme to
  identify resource on the Web}}
%

\storeglosentry{gls:xss}{name={cross-site scripting},description={Injection of HTML and /
or JavaScript into a Web page.}}%
%
\storeglosentry{gls:defacto}{
  name={de facto},
  description={A general known but undeclared fact.}
}%



